A formal approach but practical approach in train regulation

### 2025-06-23 paper/main.tex

@@

La composition est un procédé important qui nous permet de créer des chemins complexes à partir de chemins élémentaires.
Pour cela, nous décomposons cette fonction en trois parties : préparation, assemblage et nettoyage.
\\Notation : on note $o_x$ une liste d'ordres, avec $x \in \Nat$.

@@

\subsection{Nettoyage}
Cette partie fait office de "finalisation". On va d'abord s'occuper de "regrouper" les ordres des trains. Notre exemple précédent "<L,[1,2]>,<L,[3,4]>", nous donne maintenant "<L,[1,2,3,4]>". 
On ajoute également un ordre "auth" pour les "turn" utilisés par des trains pour agir sur leur propre parcours. (les trains n'utilisent pas directement "turn", mais plus simple à expliquer de cette manière)
Enfin, on chercher le premier arrêt de chaque train pour assigner les feux au rouge.
\section{Expériences}
\label{sec:experiments}

>>>>

\paragraph{Guillaume rajoute la normalisation}

%Rule application induce a relation between system state:  $S\reduces S'$ if $\gamma'$  if there is a rule for which $\gamma$ is the initial state and $\gamma'$ the conclusion. 
The reflexive-transitive closure of $\reduces$ is denoted $\reduces^*$, that is $S \reduces^* S'$ iff $S =S_0 \reduces \cdots S_n = S'$. The value $n = 0$ corresponds to $S = S'$.  A normal form of a configuration $S$ is a configuration $S'$ such that $S \reduces^* S'$ and there is no $S''$ with $S' \reduces S''$. In other words, starting from $S$, the computation can stop on $S'$. We note $S \reduces^! S'$ this relationship.

Given a system state  $S = \redTuple{x}{B}{\stateTuple{\trainSeq{}}{\regulator}{\switches}{\signals}}$ and a train $\tid{id}$, 
\begin{itemize}
\item   we say that it has no jobs if  $\trainSeq{}[\tid{id}].P = \emptyTrainProg$;
\item its source is $\mathtt{src}(S, \tid{id}) = \trainSeq{}[\tid{id}].pid$;
\item its destination is $\mathtt{dst}(S, \tid{id}) = pid \in \blocks $ if either
\begin{itemize}
\item  it has no jobs and  $\trainSeq{}[\tid{id}].pid = pid$  or
\item  the last order in  $\trainSeq{}[\tid{id}].P$ is $\su{d}{pid}$ for some direction $d$.
\end{itemize}
\end{itemize}

A system state $S = \redTuple{\guardT}{\emptyset}{\stateTuple{\trainSeq{}}{\regulator}{\switches}{\signals}} $  is \emph{reliable} if: 
\begin{itemize}
\item  no two trains meet on the same block or the same switch, \todo{En dire plus}
\item $S$ has  a unique normal form $S'$ and 
\item any train in $S'$ has no jobs and 
\item for any train $\tid{id}$, $\mathtt{dst}(S, \tid{id}) =  \mathtt{dst}(S', \tid{id})$. 
\end{itemize}

@@

The main principle behind our routing process is composition that we discuss now. A route for a train is a sequence $\vec{x} = x_1, \ldots, x_k$ of (contiguous) blocks and switches. Two routes $\vec{x}$ and $\vec{y}$ have a critical resource if  $x_i = y_j$ for some $i,j$. A \emph{global routing} is a route for each train. To compute global routing, we decompose them in  safe steps. A safe step is a global routing such that the routes share no critical resources, neither a block, neither a switch. It is clear that any global routing can be decomposed into such safe steps.

This strategy has a main advantage, it is quite flexible to new events. Suppose that at some point a train is blocked, we may recompute safe steps for the other trains, possibly releasing a deadlock. But, that being said, we are now turned to the transformation of safe steps into train orders and regulator rules.

@@

\subsection{Nettoyage}
Cette partie fait office de "finalisation". On va d'abord s'occuper de "regrouper" les ordres des trains. Notre exemple précédent "<L,[1,2]>,<L,[3,4]>", nous donne maintenant "<L,[1,2,3,4]>". 
On ajoute également un ordre "auth" pour les "turn" utilisés par des trains pour agir sur leur propre parcours. (les trains n'utilisent pas directement "turn", mais plus simple à expliquer de cette manière)
Enfin, on chercher le premier arrêt de chaque train pour assigner les feux au rouge.

\section{Expériences}
\label{sec:experiments}
